---
title: "TORC1 signaling and its role in ASR in yeasts"
author: Bin He
date: "2022-11-20 (updated `r Sys.Date()`)"
output:
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---

```{r setup, message=FALSE}
require(multcomp)
require(tidyverse)
require(cowplot)
```

## Goal
Jinye performed Western Blot to probe P-Rps6 and total Rps6 (aka S6) as a readout for the TORC1 activity. Based on the literature, active TORC1 signaling results in the phosphorylation of Rps6 (ribosomal subunit 6, indicative of active growth). Nitrogen starvation is known to suppress TORC1 and therefore leads to the loss of P-Rps6. The total Rps6 level doesn't change dramatically depending on the environmental conditions. Our goal is to use P-Rps6/totalS6 as a reporter for TORC1 activity and compare it between _S. cerevisiae_ and _C. glabrata_ under rich, no nitrogen and no phosphate conditions. Additionally, Jinye profiled two outgroup species, _K. lactis_ and _L. waltii_

## Sc vs Cg
### Data
Three replicates performed on 2021-12-17, 2022-01-11 and 2022-01-13. The blots were imaged and quantified using Li-COR software. In particular, each gel was first stained with a total protein stain, imaged, and then probed with antibodies for either P-Rps6 or total S6. The signals of the Western Blot bands were background subtracted using the regions surrounding the lanes, and normalized by the whole protein stain signal for the entire lane. The data were combined into a single text file.

```{r}
dat <- read_tsv("../input/20221120-Rps6-phospho-WB-quant-combined.txt", col_types = cols())
```

We will normalize the signals against the whole protein stain, and then further divide the normalized values between the P-Rps6 and total S6. The resulting values are proportional to the % of phosphorylated Rps6. We could normalize the two starvation treatment values to the rich condition to make the numbers a bit more interpretable.
```{r}
dt <- dat %>% 
  mutate(norm.v = signal/whole_prot, .keep = "unused") %>% # remove the variables used to create the new one
  pivot_wider(names_from = target, values_from = norm.v) %>% 
  mutate(PS6.v.tS6 = `P-Rps6` / `total Rps6`,
         species = fct_recode(species, `S. cerevisiae` = "Sc", `C. glabrata` = "Cg"),
         treatment = factor(treatment, levels = c("rich", "noNi", "noPi"), labels = c("Rich", "-N", "-Pi"))) %>% 
  arrange(species, date, treatment)
```

### Plotting
```{r}
ggplot(dt, aes(x = species, y = PS6.v.tS6, group = treatment)) +
  geom_bar(aes(fill = treatment), stat = "summary", fun = "mean", 
           position = position_dodge(0.9), width = .8, color = "gray30") +
  geom_point(position = position_dodge(0.9), size = 2, shape = 18) +
  #scale_color_manual(values = c("S. cerevisiae" = "grey20", "C. glabrata" = "orange2")) +
  scale_fill_grey(start = 0.3, end = 0.95) +
  ylab("P-Rps6 / total Rps6") +
  theme_cowplot(line_size = 0.7) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = rel(1.2), face = 3),
        legend.position = "top")
ggsave("../output/20221120-Rps6-WB-quant-compare.png", width = 3.5, height = 3.5)
```

### Statistical tests
Perform pairwise t tests to assess significance. For each species, there will be two comparisons: -Ni vs rich and -Pi vs rich. P-values will be corrected for multiple testing.

> I tried two other approaches also. 1) Wilcoxon signed-rank test or Mann-Whitney U test. These rank based tests do not assume normality and thus are more robust to violations of the normal distribution assumption. However, their lack of power is especially obvious in the case of a small sample size such as here. With a sample size of 3, the smallest p-value achivable by either test is 0.25 (4 samples goes to 0.06). 2) I also tried a linear modeling approach using aov() followed by multcomp::glht() (see code below). This approach estimates a shared error distribution from all data. I can't quite justify the shared error distribution for the two species and different conditions. Hence I eventually settled for the simple pairwise t-test.

```{r}
tmp <- dt %>% 
  mutate(
    sp = factor(species, levels = c("C. glabrata", "S. cerevisiae"), 
                 labels = c("C", "S")),
    group = factor(paste(gsub(" ", "", treatment), sp, sep = ":"),
                   levels = c("Rich:C", "-N:C", "-Pi:C",
                              "Rich:S", "-N:S", "-Pi:S"))
  ) %>% 
  select(date, species, treatment, x = PS6.v.tS6, group)
tmp %>% group_by(species, treatment) %>% 
  summarize(mean = mean(x), sd = sd(x), .groups = "drop")
```

```{r}
my_pairwise_ttest <- function(groups, dat = tmp){
  dat1 <- filter(dat, group %in% groups)
  res <- t.test(x ~ group, paired = TRUE, alternative = "t", data = dat1)
  return(broom::tidy(res) %>% select(estimate, p.value))
}
```

```{r}
myContrasts <- list(
  "Cg -N vs Rich" = c("-N:C", "Rich:C"),
  "Cg -Pi vs Rich" = c("-Pi:C", "Rich:C"),
  "Sc -N vs Rich" = c("-N:S", "Rich:S"),
  "Sc -Pi vs Rich" = c("-Pi:S", "Rich:S")
)
res <- map_dfr(myContrasts, my_pairwise_ttest, .id = "comparison") %>% 
  mutate(
    p.bonf = p.adjust(p.value, method = "bonferroni"),
    p.holm = p.adjust(p.value, method = "holm")
  )
res
```


(not in use) Set contrasts, following pp4 from https://cran.r-project.org/web/packages/multcomp/vignettes/multcomp-examples.pdf
```{r}
tmp.aov <- aov(x ~ group-1, data = tmp)
sprintf("The levels of the grouping variable are %s", 
        paste(gsub("group", "", names(coef(tmp.aov))), collapse = " "))
K <- rbind(
  "Cg -N vs rich" = c(-1, 1, 0, 0, 0, 0), 
  "Cg -Pi vs rich" = c(-1, 0, 1, 0, 0, 0), 
  "Sc -N vs rich" = c(0, 0, 0, -1, 1, 0), 
  "Sc -Pi vs rich" = c(0, 0, 0, -1, 0, 1) 
)
colnames(K) <- levels(tmp$group)
K
summary(glht(tmp.aov, linfct = K))
```

## Four species
### Data
Three replicates performed on 2022-02-16, 2022-03-08 and 2022-03-11. The blots were imaged and quantified using Li-COR software. In particular, each gel was first stained with a total protein stain, imaged, and then probed with antibodies for either P-Rps6 or total S6. The signals of the Western Blot bands were background subtracted using the regions surrounding the lanes, and normalized by the whole protein stain signal for the entire lane. The data were combined into a single text file.

```{r}
dat1 <- read_tsv("../input/20220216-Rps-phospho-WB-quant-combined-4-species.tsv", col_types = cols())
```

We will normalize the signals against the whole protein stain, and then further divide the normalized values between the P-Rps6 and total S6. The resulting values are proportional to the % of phosphorylated Rps6. We could normalize the two starvation treatment values to the rich condition to make the numbers a bit more interpretable.
```{r}
dt1 <- dat1 %>% 
  mutate(norm.v = signal/whole_prot, .keep = "unused") %>% # remove the variables used to create the new one
  pivot_wider(names_from = target, values_from = norm.v) %>% 
  mutate(PS6.v.tS6 = `P-Rps6` / `total Rps6`,
         species = factor(species,
                          levels = c("Cg", "Sc", "Kl", "Lw"),
                          labels = c("C. glabrata", "S. cerevisiae",
                                     "K. lactis", "L. waltii")),
         treatment = factor(treatment, 
                            levels = c("SC", "Nitrogen Starvation", "Pi starvation"), 
                            labels = c("Rich", "-N", "-Pi")))
```

### Plotting
```{r}
ggplot(dt1, aes(x = species, y = PS6.v.tS6, group = treatment)) +
  geom_bar(aes(fill = treatment), stat = "summary", fun = "mean", 
           position = position_dodge(0.9), width = .8, color = "gray30") +
  geom_point(position = position_dodge(0.9), size = 2, shape = 18) +
  #scale_color_manual(values = c("S. cerevisiae" = "grey20", "C. glabrata" = "orange2")) +
  scale_fill_grey(start = 0.3, end = 0.95) +
  ylab("P-Rps6 / total Rps6") +
  theme_cowplot(line_size = 0.7) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = rel(1.2), face = 3),
        legend.position = "top")
#ggsave("../output/20230105-4sps-Rps6-WB-quant-compare.png", width = 3.5, height = 3.5)
```


```{r}
ggplot(dt1, aes(x = treatment, y = PS6.v.tS6)) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.5) +
  geom_bar(aes(fill = treatment), 
           stat = "summary", fun = "mean", width = .8, color = "gray30") +
  geom_point(size = 2, color = "orange2") +
  #scale_color_manual(values = c("S. cerevisiae" = "grey20", "C. glabrata" = "orange2")) +
  scale_fill_grey(start = 0.3, end = 0.95) +
  ylab("P-Rps6 / total Rps6") +
  facet_wrap(~species, nrow = 1) +
  theme_cowplot(line_size = 0.7) +
  background_grid(major = "y") +
  theme(axis.title.x = element_blank(),
        strip.text.x = element_text(size = rel(1), face = 3),
        legend.position = "none")
ggsave("../output/20230105-4sps-Rps6-WB-quant-compare.png", width = 6, height = 3)
```

### Statistical tests
```{r}
tmp <- dt1 %>% 
  mutate(
    group = factor(paste(treatment, str_sub(species, 1, 1), sep = ":"),
                   levels = paste(rep(c("Rich", "-N", "-Pi"), times = 4),
                                  rep(c("C", "S", "K", "L"), each = 3), sep = ":"))
  ) %>% 
  select(date, species, treatment, x = PS6.v.tS6, group)
levels(tmp$group)
```

```{r}
myContrasts <- list(
  "Cg -N vs Rich" = c("-N:C", "Rich:C"),
  "Cg -Pi vs Rich" = c("-Pi:C", "Rich:C"),
  "Sc -N vs Rich" = c("-N:S", "Rich:S"),
  "Sc -Pi vs Rich" = c("-Pi:S", "Rich:S"),
  "Kl -N vs Rich" = c("-N:K", "Rich:K"),
  "Kl -Pi vs Rich" = c("-Pi:K", "Rich:K"),
  "Lw -N vs Rich" = c("-N:L", "Rich:L"),
  "Lw -Pi vs Rich" = c("-Pi:L", "Rich:L")
)
res <- map_dfr(myContrasts, my_pairwise_ttest, .id = "comparison") %>% 
  mutate(
    p.bonf = p.adjust(p.value, method = "bonferroni"),
    p.holm = p.adjust(p.value, method = "holm")
  )
res
```