---
title: "ROS_staining"
author: "JY"
date: "3/29/2023"
output: 
  html_notebook:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r}
library(tidyverse)
library(cowplot)
#library(broom)
```
load the raw data:
1) merged from d032423 to d032823 (5 biological replicates) --> merge-add_gate.csv
2) treatment info --> Sample_treatment.csv
```{r}
# import the data
bud <- read.csv("input/d042123_budding_CY.csv")

```


plot the gated data
```{r}
# plot function
unbud.timecourse <- list(
  stat_summary(fun = "mean", geom = "point", size = 3),
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 3),
  stat_smooth(aes(size = Species == "Cg"), method = "loess", 
              formula = 'y~x', se = FALSE),
  scale_x_continuous(breaks = seq(0,135,45)),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x = "Time (min)", y = "unbud (%)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.05,0.75), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)

# plot the unbuded  %
bud %>% filter(!is.na(Percent_unbud)) %>% 
  ggplot(aes(x = Time_point_min, y = Percent_unbud, color = Species)) + unbud.timecourse 
ggsave("output/20230421-bud.png", width = 8, height = 6)  

```



statistical analysis for different gate
```{r}
# Cg different gates
ros.ps %>% filter(., Gate == "500_gate" & Species == "Cg") %>% 
  lm(X.Gated ~ Treatment, data = .) %>%  # did one-way anova using different treatment
  summary()

ros.ps %>% filter(., Gate == "High_ROS" & Species == "Cg") %>% 
  lm(X.Gated ~ Treatment, data = .) %>%  # 
  summary()

ros.ps %>% filter(., Gate == "2000_gate" & Species == "Cg") %>% 
  lm(X.Gated ~ Treatment, data = .) %>%  
  summary()

# Sc with different gates
ros.ps %>% filter(., Gate == "500_gate" & Species == "Sc") %>% 
  lm(X.Gated ~ Treatment, data = .) %>%  
  summary()

ros.ps %>% filter(., Gate == "High_ROS" & Species == "Sc") %>% 
  lm(X.Gated ~ Treatment, data = .) %>%  
  summary()

ros.ps %>% filter(., Gate == "2000_gate" & Species == "Sc") %>% 
  lm(X.Gated ~ Treatment, data = .) %>%  
  summary()


```
statistical analysis for different median
```{r}
# Cg different gates
ros.ps %>% filter(., Gate == "500_gate" & Species == "Cg") %>% 
  lm(X.Median ~ Treatment, data = .) %>%  # did one-way anova using different treatment
  summary()

ros.ps %>% filter(., Gate == "High_ROS" & Species == "Cg") %>% 
  lm(X.Median ~ Treatment, data = .) %>%  # 
  summary()

ros.ps %>% filter(., Gate == "2000_gate" & Species == "Cg") %>% 
  lm(X.Median ~ Treatment, data = .) %>%  
  summary()

# Sc with different gates
ros.ps %>% filter(., Gate == "500_gate" & Species == "Sc") %>% 
  lm(X.Median ~ Treatment, data = .) %>%  
  summary()

ros.ps %>% filter(., Gate == "High_ROS" & Species == "Sc") %>% 
  lm(X.Median ~ Treatment, data = .) %>%  
  summary()

ros.ps %>% filter(., Gate == "2000_gate" & Species == "Sc") %>% 
  lm(X.Median ~ Treatment, data = .) %>%  
  summary()


```


## Goal
The goal of this set of experiments is to quantify the intracellular ROS levels to determine if phoshate starvation leads to an increased ROS level. The reason for the question is because if phoshate starvation induces ROS compared with hydrogen peroxide, and whether the changes differ between species.


## Data
old analysis chunk 
```{r}
treat <- c("Ctrl","0Pi","0.2mM","1mM","2mM","10mM","20mM","100mM")
species <- c(Cg = "C. glabrata", Sc = "S. cerevisiae")
sample <- read_csv("../input/Sample_treatment.csv", col_types = cols()) %>% 
  # arrange the factor level
  mutate(Treatment = factor(Treatment, levels = treat),
         Species = factor(Species, levels = c("Cg", "Sc"), labels = species)
  )

raw <- read_csv("../input/ROS_staining_data_raw.csv", col_types = cols())

ros <- left_join(raw, sample, by = "Sample") %>% 
  # remove a couple of H2O2 treatment levels
  #mutate(group = paste(Species, Treatment, sep = "_")) %>% 
  #filter(group %in% c("Cg_0.2mM", "Cg_1mM", ))
  select(plate = Plate, treatment = Treatment, species = Species, sample = Sample,
         count = Count, highROS = `%Gated`, x.mean = `X Mean`, x.median = `X Median`,
         x.rSD = `X rSD`, x.pCV = `X %CV`)
                            
```

```{r}
ros %>% 
  group_by(plate, treatment, species) %>% 
  summarize(across(count:median, ~ mean(.x)))
```


## Plotting
```{r}
# plot two species together
ros %>% 
  ggplot(aes(x = Treatment, y = X.Gated, color = Species)) + geom_point(size = 0.2) + 
  stat_summary(fun = "mean", geom = "point", size = 2) +
  theme_bw() +
  theme(legend.position="right", legend.box = "vertical", legend.text = element_text(size=10), 
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle=0,vjust=-0.5), axis.text = element_text(size = 12), 
        axis.title.x = element_text(vjust=-2),
        plot.title = element_text(size = 14, hjust = 0)) +
  labs(y = "Relative ROS(%)")
ggsave("../output/20230329-Raw-Cg&Sc-ROS-Staining.png", width = 8, height = 6)

# separately plot the species Cg
ros %>% filter(., Species == "Cg") %>% 
  ggplot(aes(x = Treatment, y = X.Gated)) +
  geom_boxplot(width=0.3) +
  geom_dotplot(binaxis = 'y', binwidth=0.04, stackdir = "center") +
  theme_bw() +
  theme(legend.position="right", legend.box = "vertical", legend.text = element_text(size=10), 
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle=0,vjust=-0.5), axis.text = element_text(size = 12), 
        axis.title.x = element_text(vjust=-2),
        plot.title = element_text(size = 14, hjust = 0)) +
  labs(y = "Relative ROS(%) in C. glabrata")
ggsave("../output/20230329-Cg-ROS-Staining.png", width = 8, height = 6)

# separately plot the species Sc
ros %>% filter(., Species == "Sc") %>% 
  ggplot(aes(x = Treatment, y = X.Gated)) +
  geom_boxplot(width=0.3) +
  geom_dotplot(binaxis = 'y', binwidth=0.04, stackdir = "center") +
  theme_bw() +
  theme(legend.position="right", legend.box = "vertical", legend.text = element_text(size=10), 
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle=0,vjust=-0.5), axis.text = element_text(size = 12), 
        axis.title.x = element_text(vjust=-2),
        plot.title = element_text(size = 14, hjust = 0)) +
  labs(y = "Relative ROS(%) in S. cerevisiae")
ggsave("../output/20230329-Sc-ROS-Staining.png", width = 8, height = 6)

# Draw two species side by side
ros %>%
  ggplot(aes(x = Treatment, y = X.Gated, color = Species)) +
  geom_boxplot(width=0.5) +
  geom_dotplot(binaxis = 'y', binwidth=0.04, stackdir = "center") +
  theme_bw() +
  theme(legend.position="right", legend.box = "vertical", legend.text = element_text(size=10), 
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle=0,vjust=-0.5), axis.text = element_text(size = 12), 
        axis.title.x = element_text(vjust=-2),
        plot.title = element_text(size = 14, hjust = 0.5)) +
  labs(y = "Relative ROS(%)")
ggsave("../output/20230329-Cg&Sc-ROS-Staining.png", width = 8, height = 6)

```

```{r}
# statistical analysis between groups, compare all the categorical variable (0Pi, H2O2 treatment) to Ctrl treatment
## One-way ANOVA
# Cg
ros %>% 
  filter(Species == "Cg") %>% # for a specific species
  lm(X.Gated ~ Treatment, data = .) %>%  # did one-way anova using different treatment
  summary()

ros %>% 
  filter(Species == "Sc") %>% # for a specific species
  lm(X.Gated ~ Treatment, data = .) %>%  # did one-way anova using different treatment
  summary()

#pair-wise t-test 
#(I do not know how to write the code to compare it, so I mannually generated the wide_data_format.csv from the ros file)
ros.wide <- read.csv("../input/wide_data_format.csv")
ros.wide%>% 
group_by(Treatment) %>% 
do(tidy(t.test(.$Cg, 
               .$Sc, 
               mu = 0, 
               alt = "two.sided", 
               paired = TRUE, 
               conf.level = 0.95)))
# Conlusion: no significant difference of ROS level between Cg and Sc at Ctrl, 0mM Pi and 0.2mM H2O2 treatment. 

# two way anova to include both species and treatment
ros %>% 
  filter(!Treatment %in% c("20mM","100mM")) %>% 
  aov(X.Gated ~ Species * Treatment, data = .) %>% 
  summary()

ros %>% 
  filter(!Treatment %in% c("20mM","100mM")) %>% 
  aov(X.Gated ~ Species * Treatment, data = .) %>%
  TukeyHSD(., which = "Treatment")
```
