---
title: "CTA1 mRNA and protein induction (Figure 3)"
author: "Bin He"
date: "2022-10-25 (updated `r Sys.Date()`)"
output: 
  html_notebook:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r}
library(tidyverse)
library(ggtext)
library(cowplot)
```


# _CTA1_ mRNA induction [Fig. 3A]
Experiment was performed on 2020/03/25-27, see [here](https://docs.google.com/spreadsheets/d/1dClNE3oDJyQQlZwcf_NoHs56Wi1G938K44yyCrwCpCQ/edit?usp=sharing) for experimental methods. Briefly, _C. glabrata_ cells were grown to mid-log phase (OD~0.5), washed and treated with either SC media (mock), 0mM Pi (no Pi) or 10 mM H2O2 for 45 minutes. RNA was extracted and qRT-PCR was performed with _ACT1_ used as a reference gene and the mock treated sample as the reference sample. The following dataset was extracted from the above spreadsheet.
```{r}
rna <- read_tsv("input/CTA1-mRNA-qPCR/20200327-CTA1-mRNA-Ct.txt") %>% 
  mutate(Treatment = fct_recode(Treatment, 
                                H2O2 = "10mM H2O2", 
                                `-Pi` = "no Pi", 
                                Mock = "untreated"))
# check the control wells
rna %>% filter(Group %in% c("noRT", "NTC"))
# get the subset for "unknowns" vs "standards"
rnaT <- rna %>% filter(Group == "unknown")
rnaS <- rna %>% filter(Group == "standard")
```
> C10 and D12 are problematic

Check the standard curve
```{r}
p <- rnaS %>% 
  ggplot(aes(x = log10(Dilution), y = Cp)) + 
  geom_point(aes(color = Target)) +
  scale_color_brewer(palette = "Set1") +
  theme_cowplot()
p
```
```{r}
rnaS %>% 
  #select(Target, Dilution, Cp) %>% 
  mutate(x = log10(Dilution)) %>% 
  group_by(Target) %>% 
  nest() %>% 
  mutate(test = map(data, function(df) lm(Cp ~ x, data = df)),
         slope = map(test, tidy)) %>% 
  unnest(slope) %>% 
  filter(term == "x") %>% 
  select(Target, slope = estimate, std.error, p.value)
  
```

> primer set C behaves better than primer set A

Visualize the Ct values for the unknown samples
```{r}
p <- rnaT %>% 
  ggplot(aes(x = Treatment, y = Cp)) + 
  geom_point(aes(color = Name), shape = 16, size = 2.5, position = position_dodge(width = 0.4)) + 
  facet_wrap(~Target) +
  theme_bw(base_size = 16)
p
```

Calculate ∆∆Ct. Because the dilution amount is the same across all samples, we don't have to make any adjustments. `r all(rnaT$Dilution == 100)`
```{r}
rna.dCt <- rnaT %>% 
  # calculate mean Ct for the reference gene
  filter(Target == "ACT1") %>% 
  group_by(Name, Treatment) %>% 
  summarize(refG = mean(Cp), .groups = "drop") %>% 
  # append the reference gene Ct mean to the original data
  right_join(filter(rnaT, Target != "ACT1")) %>% 
  # subtract the reference gene Ct mean from each target
  mutate(dCt = Cp - refG) %>% 
  select(-refG)

rna.ddCt <- rna.dCt %>% 
  # calculate mean dCt for the reference sample
  filter(Treatment == "Mock") %>% 
  group_by(Name, Target) %>% 
  summarize(refT = mean(dCt), .groups = "drop") %>% 
  right_join(rna.dCt) %>% 
  mutate(ddCt = dCt - refT, FC = 2^-ddCt) %>% 
  select(-refT)
```

Visualize fold changes relative to the mock treated sample and normalized by ACT1. Using CTA1 primer set A. Overall pattern is highly consistent. Primer set C data contains one outlier that is most likely due to experimental mistakes -- one -Pi sample showed >400 fold induction 
```{r}
treat_color <- c(`H<sub>2</sub>O<sub>2</sub>` = "#E69138", `-Pi` = "#9900FF", Mock = "#666666")
p.rna <- filter(rna.ddCt, Target == "CTA1_A") %>% 
  mutate(Strain = fct_recode(Name, `wild type 1` = "yH001", `wild type 2` = "yH002"),
         Treatment = fct_recode(Treatment, `H<sub>2</sub>O<sub>2</sub>` = "H2O2")) %>% 
  ggplot(aes(x = Treatment, y = FC)) +
  geom_point(aes(shape = Strain, color = Treatment), size = 3.5, alpha = 0.9,
             position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.2)) +
  scale_shape_discrete(solid = TRUE) +
  scale_color_manual(values = treat_color, guide = "none") +
  stat_summary(aes(group = Strain), fun = "mean", geom = "crossbar", color = "darkred",
               position = position_dodge(0.6), width = 0.4) +
  scale_y_log10() + ylab("CTA1 mRNA over mock") +
  theme_cowplot(line_size = 0.7, font_size = 16) +
  theme(legend.text = element_text(size = rel(0.8)),
        legend.title = element_text(size = rel(0.9)),
        legend.position = "top",
        axis.title.x = element_blank(),
        axis.text.x = element_markdown(size = rel(1.1)),
        axis.text.y = element_text(size = rel(1.1)))
p.rna
ggsave("output/20221213-CTA1-mRNA-qRT-PCR.png", width = 4, height = 3.5)
```

Statistical tests
```{r}
lm <- rna.ddCt %>% filter(Target == "CTA1_A") %>% 
  mutate(Treatment = fct_relevel(Treatment, "Mock")) %>% 
  rename(Strain = Name) %>% 
  lm(ddCt ~ Treatment + Strain, data = .)
summary(lm)
```

# Cta1-GFP protein levels [Fig. 3B]
## Common plotting functions
**Update 2022-11-08** 
This list is used as a shared plotting set up.
```{r}
p.timecourse <- list(
  stat_summary(fun = "mean", geom = "point", size = 3),
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 3),
  stat_smooth(method = "loess", formula = 'y~x', se = FALSE),
  scale_x_continuous(breaks = seq(0,240,60)),
  scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.05,0.75), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)
```

## Prepare data
```{r}
pi.dat.long <- read_csv("input/wt-Pi-H2O2/Long_data_CTA1-GFP.csv", col_types = cols()) %>% 
  mutate(new_time = as.numeric(gsub("min","",Time))) %>% 
  arrange(new_time, Treatment)
```

### Bin's version
```{r}
treat_color <- c(`2mM H2O2` = "#E69138", `0mM Pi` = "#9900FF", Control = "#666666")

p.gfp <- pi.dat.long %>%
  filter(!is.na(median)) %>% 
  ggplot(aes(x = new_time, y = median, color = Treatment)) +
  p.timecourse +
  scale_color_manual(values= treat_color, labels = c("-Pi", bquote(~H[2]*O[2]), "Mock")) +
  theme(legend.text = element_text(size = rel(0.85), face = 1),
        legend.position="top",
        legend.title = element_blank())
plot_grid(p.rna, NULL, p.gfp, nrow = 1, rel_widths = c(1, 0.1, 1), align = "hv")
#ggsave("output/20221117-Pi-H2O2-timecourse.png", width = 4, height = 3.5)
ggsave("output/20221214-CTA1-rna-gfp-align.png", width = 8, height = 3.5)
```
