---
title: "MET3pr-CTA1 induction, validate (Fig. S5)"
author: "Bin - strain, analysis - Lindsey, flow;"
date: "2023-09-01 (updated `r Sys.Date()`)"
output: 
  html_notebook:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r}
library(tidyverse)
library(ggtext)
library(cowplot)
```

## Goal
Validate _MET3pr-CTA1_ induction under -Met -Cys media condition

## Data import and prep
```{r}
tp <- c("0",  "30", "45", "60", "90", "120") # ordered time points

raw <- read_tsv("input/MET3pr-CTA1/20230901-METpr-CTA1-induction-lfs.tsv",
                col_types = cols(), na = c("", "NA", "N/A"), comment = "#") %>% 
  mutate(Time = ordered(time, levels = tp))
```

merge with sample metadata
```{r}
sample <- tribble(
  ~strain, ~genotype,
  "yH298", "CTA1pr-CTA1",
  "yH299", "CTA1pr-CTA1",
  "yH001", "WT noGFP",
  "yH181", "WT noGFP",
  "yH347", "cta1pr∆-CTA1",
  "met3-1", "MET3pr-CTA1",
  "met3-2", "MET3pr-CTA1",
  "met3-3", "MET3pr-CTA1",
  "met3-4", "MET3pr-CTA1"
)

gt <- c("WT noGFP", "cta1pr∆-CTA1", "CTA1pr-CTA1", "MET3pr-CTA1")

# combine genotype and treat variables into the flow data
dat <- left_join(raw, sample, by = c("strain")) %>% 
  mutate(genotype = factor(genotype, levels = gt),
         treatment = factor(treatment, levels = c("Pi", "mock", "met"),
                            labels = c("-Pi", "Rich", "-MC")))

# experiment structure
dat %>% 
  filter(time == 0) %>% 
  group_by(genotype, treatment, day) %>% 
  summarize(nEx = n()) %>% 
  pivot_wider(names_from = day, values_from = nEx, names_prefix = "D")
```


## Plotting
### Select data for plotting
For the figure, we will show the _CTA1pr-CTA1-GFP_ and _MET3pr-CTA1-GFP_ strains under either their inducing condition or mock treatment. We will use Day 1 data as the representative
```{r}
dat.plot <- dat %>% 
  mutate(group = paste(genotype, treatment, sep = " : ")) %>% 
  filter(group %in% c("CTA1pr-CTA1 : -Pi", #"CTA1pr-CTA1 : Rich",
                      "MET3pr-CTA1 : -MC"))#, "MET3pr-CTA1 : Rich"))
```

### Common plotting functions
**Update 2022-11-08** 
This list is used as a shared plotting set up.
```{r}
p.timecourse <- list(
  stat_summary(fun = "mean", geom = "point", size = 3),
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 3),
  stat_smooth(method = "loess", formula = 'y~x', se = FALSE),
  scale_x_continuous(breaks = seq(0,240,30)),
  #scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x = "Time (min)", y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.05,0.72), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)
```

### Plot
```{r}
dat.plot %>%
  mutate(median = median / 1000) %>% 
  ggplot(aes(x = time, y = median, color = group)) +# p.timecourse +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 3) +
  stat_smooth(method = "loess", formula = 'y~x', se = FALSE) +
  scale_x_continuous(breaks = c(0,30,45,60,90,120)) +
  scale_color_manual("Strain:treatment", values = c("#9966DD", "#5DC863")) +
  labs(x = "Time (min)", y = "Cta1-GFP protein level (a.u.)") +
  theme_cowplot(line_size = 0.7, font_size = 14) +
  theme(legend.position=c(0.05,0.85), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )

ggsave("output/20230904-MET3pr-CTA1-induction-timecourse.png", width = 3.5, height = 4)
```

