---
title: "Effect of rim15∆ effects on CTA1 induction (Figure 6)"
author: "HB"
date: "2023-05-17 (updated `r Sys.Date()`)"
output: 
  html_notebook:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r}
library(tidyverse)
library(ggtext)
library(cowplot)
```

## Goal
We hypothesize that Rim15 is upstream of Msn4 and contributes to _CTA1_ induction during -Pi. Our approach is to delete _RIM15_ and compare Cta1-GFP induction in this background with that under a wild type background.

## Data import and prep
```{r}
tp <- c("0",  "30", "60", "90", "120", "150",  "180", "210", "240") # ordered time points

raw <- dir("input/rim15_gz_merge", pattern = "merge", full.names = TRUE) %>% 
  map_dfr(read_csv, col_types = cols(), na = c("", "NA", "N/A"), .id = "replicate") %>% 
  filter(`X Parameter` == "BL1-H", is.na(`Y Parameter`)) %>% 
  select(replicate, plate = Plate, Group, well = Sample, 
         count = Count, median = `X Median`,  peak = `X Peak`, 
         sd = `X SD`, cv = `X %CV`, rCV = `X %rCV`) %>% 
  mutate(Time = ordered(Group, levels = paste0(tp, "min"), labels = tp),
         time = as.numeric(gsub("min", "", Group)),
         Group = NULL) %>% 
  arrange(time)
```

merge with sample metadata
```{r}
sample <- dir("input/rim15_treatment", pattern = "^d", full.names = TRUE) %>% 
  map_dfr(read_csv, na = c("","NA","N/A"), col_types = cols(), .id = "replicate") %>% 
  filter(!is.na(group_name)) %>% 
  select(replicate, genotype = group_name, well = Sample, treat) %>% 
  mutate(treat = fct_recode(treat, `-Pi` = "0Pi", mock = "ctrl"))

# combine genotype and treat variables into the flow data
data <- left_join(raw, sample, by = c("replicate", "well")) %>% 
  separate(well, into = c("row", "column"), sep = 1)

# any unannotated rows?
data %>% filter(is.na(genotype) | is.na(treat))
```

filter for 0Pi data
```{r}
levels = c(WT = "wt", `msn4-` = "msn4_nat_ko", `rim15-` = "rim15_ko", `msn4- rim15-` = "msn4rim15_ko")
pi.dat <- data %>% 
  filter(treat == "-Pi" & genotype %in% c("wt","msn4_nat_ko","rim15_ko","msn4rim15_ko")) %>%  
  select(Time, plate, replicate, genotype, median) %>%
  mutate(Genotype = fct_recode(genotype, !!!levels),
         new_median = median/1000,
         new_time = as.numeric(gsub("min","",Time))) %>% 
  #mutate(Genotype = gsub("_nat","",Genotype)) %>% 
  arrange(new_time, genotype, replicate) 
```

## Plotting

### Common plotting functions
**Update 2022-11-08** 
This list is used as a shared plotting set up.
```{r}
p.timecourse <- list(
  stat_summary(fun = "mean", geom = "point", size = 3),
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 3),
  stat_smooth(method = "loess", formula = 'y~x', se = FALSE),
  scale_x_continuous(breaks = seq(0,240,30)),
  #scale_size_manual(values = c(0.8, 1.5), guide = "none"),
  labs(x = "Time (min)", y = "Cta1-GFP protein level (a.u.)"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(legend.position=c(0.05,0.72), 
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        #axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )
)
```

### Plot
```{r}
genotype_color <- RColorBrewer::brewer.pal(8, "Dark2")[c(8, 3, 7, 5, 2, 1)] 
names(genotype_color) <- c("WT", "msn4-", "rim15-", "msn4- rim15-")

pi.dat %>%
  filter(!is.na(new_median)) %>% 
  ggplot(aes(x = new_time, y = new_median, color = Genotype)) + p.timecourse +
  scale_color_manual(values= genotype_color) +
  xlab("Phosphate starvation, time (min)") +
  theme(legend.position=c(0.05,0.75),
        #legend.box.background = element_rect(color = "black"),
        #legend.box.margin = margin(3,3,3,3),
        legend.text = element_text(face = 3), 
        legend.title = element_text(),
        axis.title = element_text(size = rel(1)), 
        axis.text.x = element_text(hjust=0.5), axis.text = element_text(size = rel(1)),
  )

ggsave("output/20221118-rim15-deletion-timecourse.png", width = 4.5, height = 4)
```

