---
title: "Plot ASR pheontypic results"
author: Bin He
date: "2023-02-05 (updated `r Sys.Date()`)"
output:
  html_notebook:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: hide
---

```{r setup, message=FALSE}
require(tidyverse)
require(cowplot)
```

## Introduction
### Goal

Plotting ASR data for Fig. 1, S1, ...

### Data
Jinye's table goes here

Read in data
```{r}
raw <- read_csv("../input/Sc_Cg_Rapamycin_0Nitrogen_ASR_raw.csv", col_types = cols(), comment = "#") %>% 
  mutate(Date = gsub("d(\\d\\d)(\\d\\d)(\\d\\d)", "\\1/\\2/\\3", Date)) 
  # %>% select(-`MO/MM`, -`PO/PM`)

# data sanity check, quick view
sapply(select(raw, Species, Strain, Genotype), unique)

# use the ASR score from the raw data
asr_score <- raw %>% filter(!(ASR_Score == "NA")) %>% select(-`MO/MM`)

```

## Basal H2O2 resistance
_**Goal**_

- Compare the basal survival rates of the two species at different [H2O2] to identify comparable concentrations for ASR.

_**Experiment**_

- Jinye measured CFU for _S. cerevisiae_ and _C. glabrata_ exposed to a range of [H2O2]

_**Data**_

- Filter the data for the relevant experiments.

**Main dataset**: 

Species         H2O2                        Description
--------------  -------------               ---------------
C. glabrata     0, 40, 60mM           MO, RO, NO, no phosphate starvation as primary stress
S. cerevisiae   0, 4, 6mM             MO,RO, NO, no phosphate starvation as primary stress

Date       Species   Replicate
--------   --------  ----------
03/21/22   Sc, Cg    1


Statistical test for difference between species at the highest concentration
```{r}
tmp <- raw %>% 
  filter(H2O2 %in% c("60mM", "6mM"), Group == "MO", !(Date %in% c("01/12/23", "01/18/23", "02/02/23")))
tmp %>% group_by(Species) %>% summarize(mean = mean(`MO/MM`))

wilcox.test(`MO/MM` ~ Species, data = tmp, paired = TRUE)

tmp_4 <- raw %>% 
  filter(H2O2 %in% c("40mM", "4mM"), Group == "MO", !(Date %in% c("02/02/23")))
tmp_4 %>% group_by(Species) %>% summarize(mean = mean(`MO/MM`))

wilcox.test(`MO/MM` ~ Species, data = tmp_4, paired = TRUE)
```

## ASR at different Primary stress treatment
_**Goal**_

- Generalize the main figure finding of a stronger ASR in _C. glabrata_ than in _S. cerevisiae_ at a single primary stress treatment length by extending the analysis to multiple length of primary treatment

_**Data**_

**Main dataset**: 

Species       H2O2                       Description 
--------      -----                      ------------
C. glabrata   0, 40, 60mM             full ASR experiment
S. cerevisiae 0, 4, 6mM               full ASR experiment

Date       Species   Strain    Replicate
-----      --------  -------   ----------
07/16/22   Cg            
       

_**Plotting**_

```{r}
# expand the species names
species.label <- c(
  "Sc" = "S. cerevisiae",
  "Cg" = "C. glabrata"
)


p_Cg <- asr_score %>% 
  #1st_Stress is a character variable before converting to factor
  filter(`1st_Stress` %in% c("62.5","125"), Species == "Cg") %>% 
  mutate(`1st_Stress` = factor(`1st_Stress`,levels = c("62.5", "125"))) %>% 
  ggplot(aes(x = `1st_Stress`, y = ASR_Score)) +
  geom_hline(yintercept = 1, linetype = 2, color = "gray50") +
  geom_point(position = position_jitter(width = 0.1), size = 2) + 
  stat_summary(position = position_nudge(x = 0.2),
               fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  xlab("Primary Stress (Rapamycin ng/ml)") + ylab("ASR score") +
  facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)) +
  theme_cowplot(line_size = 0.7) +
  theme(legend.text = element_text(size = rel(1), face = 3),
        legend.position = c(0.1, 0.85),
        strip.text = element_text(size = rel(1), face = 3))
p_Cg
ggsave("../output/20230205-ASR-Rapa-Cg-all-H2O2conc.png", width = 4, height = 5)

p <- asr_score %>% 
  #1st_Stress is a character variable before converting to factor
  filter(`1st_Stress` %in% c("62.5","125")) %>% 
  mutate(`1st_Stress` = factor(`1st_Stress`,levels = c("62.5", "125"))) %>% 
  ggplot(aes(x = `1st_Stress`, y = ASR_Score)) +
  geom_hline(yintercept = 1, linetype = 2, color = "gray50") +
  geom_point(position = position_jitter(width = 0.1), size = 2) + 
  stat_summary(position = position_nudge(x = 0.2),
               fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  xlab("Primary Stress (Rapamycin ng/ml)") + ylab("ASR score") +
  facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)) +
  theme_cowplot(line_size = 0.7) +
  theme(legend.text = element_text(size = rel(1), face = 3),
        legend.position = c(0.1, 0.85),
        strip.text = element_text(size = rel(1), face = 3))
p
ggsave("../output/20230205-ASR-Rapa-species-all-H2O2conc.png", width = 4, height = 5)


p <- asr_score %>% 
  #1st_Stress is a character variable before converting to factor
  mutate(`1st_Stress` = factor(`1st_Stress`,levels = c("62.5", "125", "0Ni"))) %>% 
  ggplot(aes(x = `1st_Stress`, y = ASR_Score)) +
  geom_hline(yintercept = 1, linetype = 2, color = "gray50") +
  geom_point(position = position_jitter(width = 0.1), size = 2) + 
  stat_summary(position = position_nudge(x = 0.2),
               fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  xlab("Type of Primary Stress_combine_46") + ylab("ASR score") +
  facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)) +
  theme_cowplot(line_size = 0.7) +
  theme(legend.text = element_text(size = rel(1), face = 3),
        legend.position = c(0.1, 0.85),
        strip.text = element_text(size = rel(1), face = 3))
p
#ggsave("../output/20230205-ASR-Rapa-0Nitrogen-species-all-H2O2conc.png", width = 5, height = 3)

p_4 <- asr_score %>% 
  filter(H2O2 %in% c("40mM", "4mM")) %>% 
  #1st_Stress is a character variable before converting to factor
  mutate(`1st_Stress` = factor(`1st_Stress`,levels = c("62.5", "125", "0Ni"))) %>% 
  ggplot(aes(x = `1st_Stress`, y = ASR_Score)) +
  geom_hline(yintercept = 1, linetype = 2, color = "gray50") +
  geom_point(position = position_jitter(width = 0.1), size = 2) + 
  stat_summary(position = position_nudge(x = 0.2),
               fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  xlab("Type of Primary Stress; H2O2 at 40mM vs 4mM") + ylab("ASR score") +
  facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)) +
  theme_cowplot(line_size = 0.7) +
  theme(legend.text = element_text(size = rel(1), face = 3),
        legend.position = c(0.1, 0.85),
        strip.text = element_text(size = rel(1), face = 3))
p_4
#ggsave("../output/20230205-ASR-Rapa-0Nitrogen-species-40mMns4mM.png", width = 5, height = 3)

p_6 <- asr_score %>% 
  filter(H2O2 %in% c("60mM", "6mM")) %>% 
  #1st_Stress is a character variable before converting to factor
  mutate(`1st_Stress` = factor(`1st_Stress`,levels = c("62.5", "125", "0Ni"))) %>% 
  ggplot(aes(x = `1st_Stress`, y = ASR_Score)) +
  geom_hline(yintercept = 1, linetype = 2, color = "gray50") +
  geom_point(position = position_jitter(width = 0.1), size = 2) + 
  stat_summary(position = position_nudge(x = 0.2),
               fun.data = "mean_cl_boot", geom = "pointrange", color = "red") +
  xlab("Type of Primary Stress; H2O2 at 60mM vs 6mM") + ylab("ASR score") +
  facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)) +
  theme_cowplot(line_size = 0.7) +
  theme(legend.text = element_text(size = rel(1), face = 3),
        legend.position = c(0.1, 0.85),
        strip.text = element_text(size = rel(1), face = 3))
p_6
#ggsave("../output/20230205-ASR-Rapa-0Nitrogen-species-all-60mMvs6Mm.png", width = 5, height = 3)
```


```{r}
p_sep <- asr_score %>% 
  #filter(!is.na(ASR_score)) %>% 
  mutate(`1st_Stress` = factor(`1st_Stress`, levels = c("62.5", "125", "0Ni"))) %>% 
  ggplot(aes(x = Species, y = ASR_Score)) +
  geom_point(#aes(fill = low_count),
    position = position_jitter(0.1), size = 1.5, shape = 21) + 
  stat_summary(
               fun.data = "mean_se", geom = "pointrange", color = "red") +
  #scale_y_log10() +
  scale_shape_manual(values = c(21,22), labels = species.label) +
  #scale_fill_manual(values = c("white", "gray40"), guide = "none") +
  #facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)) +
  facet_wrap(~ `1st_Stress`, scales = "free") +
  xlab("Type of Primary Stress_combine all H2O2 conc.") + ylab("ASR score (r'/r)") +
  theme_cowplot(line_size = 0.7)
p_sep
#ggsave("../output/20230205-ASR-score-Rapa-Ni-across-1st-Stress-sep-panel-combine-46.png", width = 6, height = 3)


p_sep_4 <- asr_score %>% 
  filter(H2O2 %in% c("40mM","4mM")) %>% 
  mutate(`1st_Stress` = factor(`1st_Stress`, levels = c("62.5", "125", "0Ni"))) %>% 
  ggplot(aes(x = Species, y = ASR_Score)) +
  geom_point(#aes(fill = low_count),
    position = position_jitter(0.1), size = 1.5, shape = 21) + 
  stat_summary(
               fun.data = "mean_se", geom = "pointrange", color = "red") +
  #scale_y_log10() +
  scale_shape_manual(values = c(21,22), labels = species.label) +
  #scale_fill_manual(values = c("white", "gray40"), guide = "none") +
  #facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)) +
  facet_wrap(~ `1st_Stress`, scales = "free") +
  xlab("Type of Primary Stress; H2O2 at 40mM vs 4mM") + ylab("ASR score (r'/r)") +
  theme_cowplot(line_size = 0.7)
p_sep_4
#ggsave("../output/20230205-ASR-score-Rapa-Ni-across-1st-Stress-sep-panel-40vs4.png", width = 6, height = 3)

p_sep_6 <- asr_score %>% 
  filter(H2O2 %in% c("60mM","6mM")) %>% 
  mutate(`1st_Stress` = factor(`1st_Stress`, levels = c("62.5", "125", "0Ni"))) %>% 
  ggplot(aes(x = Species, y = ASR_Score)) +
  geom_point(#aes(fill = low_count),
    position = position_jitter(0.1), size = 1.5, shape = 21) + 
  stat_summary(
               fun.data = "mean_se", geom = "pointrange", color = "red") +
  #scale_y_log10() +
  scale_shape_manual(values = c(21,22), labels = species.label) +
  #scale_fill_manual(values = c("white", "gray40"), guide = "none") +
  #facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)) +
  facet_wrap(~ `1st_Stress`, scales = "free") +
  xlab("Type of Primary Stress; H2O2 at 60mM vs 6mM") + ylab("ASR score (r'/r)") +
  theme_cowplot(line_size = 0.7)
p_sep_6
#ggsave("../output/20230205-ASR-score-Rapa-Ni-across-1st-Stress-sep-panel-60vs6.png", width = 6, height = 3)
```