---
title: "ROS_staining"
author: "JY"
date: "3/29/2023"
output: 
  html_notebook:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r}
library(tidyverse)
library(cowplot)
```
## Goal
Chronic phosphate limitation (>5hr) has been shown to induce cell cycle arrest, manifested as a steady decrease in the percent of cells with buds in _S. cerevisiae_. This coincides with the inductioin of a "general starvation program" that is associated with large changes in metabolism and upregulation of stress responses, including a number of OSR genes (but do not include _CTT1_ and _CTA1_). Thus, one hypothesis for the divergent ASR phenotype we observed is that _C. glabrata_ enters cell cycle arrest faster than _S. cerevisiae_ and the ASR effect we observed is in fact the manifestation of its general starvation program. To test this hypothesis, Chris Youngstrom in the lab did a time course microsopy to quantify the budding index (% of cells with buds) in both species during phosphate starvation.

## Data
The experiments were performed on 2023/04/07, 04/10 and 04/11. Each "replicate" is an independent sample, for which multiple fields of view were counted on a hemotocytomer. We will first sum up the numbers for each replicate and then calculate the percentages. We will ignore the Excel calculated percentages, which were based on individual fields of views.

```{r}
# import the data
raw <- read_csv("input/20230421-budding-data-clean-HB.csv", col_types = "ccciiii") %>% 
  mutate(exp = paste(date, replicate, sep = "-"))
raw %>% group_by(species, exp, time) %>%
  summarize(fileds = n(), total_count = sum(total)) %>% 
  arrange(species, exp, time)
```

check to make sure that the total number of budded and unbudded cells add up to the total cell count
```{r}
with(raw, all(budded + unbudded == total))
```

sum up the counts from different (mostly 4) fields of view for each sample

```{r}
bud <- raw %>% 
  group_by(date, species, exp, replicate, time) %>% 
  summarize(across(c(total, budded, unbudded), sum), .groups = "drop")
bud
```

In the 04/07 experiment, Chris used the same 0 min sample for both biological replicates -- the cultures were split after taking the 0 minute time point. We will duplicate the 0 minute time point data from R1 and use it for R2. Let's also calculate the percent unbudded and the [binomial proportion confidence interval](https://en.wikipedia.org/wiki/Binomial_proportion_confidence_interval)
```{r}
subRefDat <- filter(bud, date == "d040723", time == 0) %>% 
  mutate(exp = gsub("R1", "R2", exp))
bud1 <- rbind(bud, subRefDat) %>% arrange(species, exp, time) %>% 
  mutate(p_unbud = unbudded / total,
         sd = sqrt((p_unbud * (1-p_unbud))/total))
bud1 %>% head()
```

## Plotting
04/07 experiment used different time points and didn't have the 0 time point for both replicates. Here we focus on the 04/10 and 04/11 data, although the same trend was also observed for the 04/07 data.
```{r}
sps <- c("Sc" = "S. cerevisiae", "Cg" = "C. glabrata")

bud1 %>% 
  filter(date != "d040723") %>% 
  #mutate(time = factor(time, levels = sort(unique(time)))) %>% 
  ggplot(aes(x = time, y = p_unbud, group = exp, color = replicate)) + 
  geom_point() + geom_line() +
  geom_errorbar(aes(ymin = p_unbud - 1.96*sd, ymax = p_unbud + 1.96*sd,
                    color = replicate), width = 1.5, alpha = 0.7) +
  scale_color_manual(values = c("black", "darkred"), guide = "none") +
  scale_x_continuous(breaks = c(0, 15, 45, 75, 105, 135)) +
  facet_wrap(~species, labeller = labeller(species = sps)) +
  xlab("Time after starvation (min)") + ylab("Percent unbudded") +
  theme_bw(base_size = 16) +
  theme(strip.text = element_text(face = 4))
ggsave("output/20230422-HB-unbudded-timecourse.png", width = 6, height = 4)  

```

Statistical test for 45 min data
```{r}
dat <- filter(bud1, time == 45, date != "d040723") %>% 
  group_by(species) %>% 
  summarize(across(budded:unbudded, sum)) %>% 
  column_to_rownames("species") %>% 
  as.matrix()
fisher.test(dat)
```
> In fact, _S. cerevisiae_ has a higher percent of unbudded cells compared with _C. glabrata
