---
title: "Msn4 nuclear localization under various stress and in two species"
author: Bin He
date: "2023-03-22 (updated `r Sys.Date()`)"
output:
  html_notebook:
    toc: true
    toc_depth: 4
    code_folding: hide
---

```{r}
require(tidyverse)
require(cowplot)
```

## Goal
Use fluorescent microscopy to track Msn4 (and Msn2) nuclear localization in both _S. cerevisiae_ and _C. glabrata_ under different conditions.

## _rim15∆_ effect on CgMsn4<sup>nuc</sup>

### data

| Date_from | Date_to | Strains | Experiment |
|:----------|:----------|:----------|:----------|
| 2022/11/15 | 2022/11/17 | wt, rim15∆ | determine Msn4<sup>nuc</sup> under -Pi |

```{r}
dat.f7 <- read_tsv("input/20221117-rim15-noPi-raw.tsv", col_types = cols())
```

### summary and stat test
1. sum the total number of nuc and cyto cells for the same strain across replicates
1. calculate the percentage of Msn4<sup>nuc</sup> cells in each genotype
1. conduct Fisher's exact test to determine if the difference between strains under -Pi or rich conditions are significant
```{r}
sum.f7 <- dat.f7 %>% 
  group_by(treatment, genotype) %>% 
  summarize(across(nuc:cyto, sum), .groups = "drop") %>% 
  mutate(percent = scales::percent(nuc / (nuc+cyto)))
sum.f7
```
Fisher's exact test comparing _rim15∆_ vs wild type under 0Pi condition
```{r}
tmp <- sum.f7 %>% 
  filter(treatment == "0Pi") %>% 
  select(genotype, nuc, cyto) %>% 
  column_to_rownames(var = "genotype") %>% 
  as.matrix()
fisher.test(tmp)
```


Fisher's exact test comparing _rim15∆_ vs wild type under rich media condition
```{r}
tmp <- sum.f7 %>% 
  filter(treatment == "Ctrl") %>% 
  select(genotype, nuc, cyto) %>% 
  column_to_rownames(var = "genotype") %>% 
  as.matrix()
fisher.test(tmp)
```

> The difference under 0Pi is significant, but not under rich media condition.


## Msn4<sup>nuc</sup> in _S. cerevisiae_ vs _C. glabrata_ under different stresses

### data

```{r}
dat.f5 <- read_tsv("input/20230322-wildtype-multiple-conditions-raw.tsv", col_types = cols())
dat.f5 %>% group_by(species, treatment) %>% 
  summarize(n_days = length(unique(date)), 
            n_pics = length(unique(picture_name)), 
            total_cells = sum(total_fluorescent_cell), 
            .groups = "drop") %>% 
  arrange(treatment, species)
```

### summary and stat test
1. sum the total number of nuc and cyto cells for the same strain across replicates
1. calculate the percentage of Msn4<sup>nuc</sup> cells in each genotype
1. conduct Fisher's exact test to determine if the difference between strains under -Pi or rich conditions are significant
```{r}
sum.f5 <- dat.f5 %>% 
  group_by(treatment, species) %>% 
  summarize(across(nuc:cyto, sum), .groups = "drop") %>% 
  mutate(percent = scales::percent(nuc / (nuc+cyto)))
sum.f5
```
Fisher's exact test comparing _rim15∆_ vs wild type under 0Pi condition
```{r}
my_fisher_test <- function(data, treat){
  #sprintf("Testing %s", treat)
  tmp <- data %>% 
    filter(treatment == treat) %>% 
    select(species, nuc, cyto) %>% 
    column_to_rownames(var = "species") %>% 
    as.matrix()
  fisher.test(tmp)
}
```

```{r}
treatments <- setdiff(sum.f5$treatment, "H2O2")
FET.res <- sapply(treatments, function(x) my_fisher_test(sum.f5, x)) %>% 
  t() %>% as.data.frame() %>% 
  rownames_to_column(var = "treatment") %>% 
  as_tibble() %>% 
  select(treatment, p.value, odds.ratio = estimate) %>% 
  unnest(p.value:odds.ratio) %>% 
  mutate(p.adj = p.adjust(p.value, method = "bonf"))
FET.res %>% select(treatment, odds.ratio, p.value, p.adj) %>% arrange(desc(odds.ratio))
```

