---
title: "Bin_code_adpt"
author: "Jinye Liang"
date: "Sep 13, 2023"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(mgcv)
library(broom)
library(viridis)
```


### Common plotting functions

```{r}
species.label <- c(Sc = "S. cerevisiae", Cg = "C. glabrata")

p.survival <- list(
  geom_point(shape = 1, stroke = 1, size = 2, 
             position = position_jitter(width = 0.15)),
  stat_summary(fun = mean, fun.max = mean, fun.min = mean,
               geom = "crossbar", color = "red", width = 0.5),
  facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)),
  scale_y_continuous(labels = scales::percent),
  ylab("% survival"),
  theme_cowplot(line_size = 0.7, font_size = 14),
  theme(strip.text = element_text(size = rel(1), face = 3))
)

p.asr <- list(
  geom_hline(yintercept = 1, linetype = 2, color = "gray50"),
  geom_point(position = position_jitter(width = 0.1), size = 2),
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange", color = "red",
               position = position_nudge(x = 0.2)),
  facet_wrap(~Species, scales = "free_x", labeller = as_labeller(species.label)),
  theme_cowplot(line_size = 0.7),
  theme(strip.text = element_text(size = rel(1.1), face = 3))
)
```


## Supplementary Fig: ASR in msn4∆ and skn7∆ vs wt
_**Goal**_

Determine if transcriptional factor (TF) msn4 and skn7 contribute to the ASR.

_**Rationale**_

Msn4 and Skn7 are stress responsive TFs regulating the Cta1-GFP induction during phosphate starvation. The reviewer is wondering if the loss of Msn4 or Skn7 affect the ASR for H2O2. To answer this quesion, Jinye performed the CFU assay for H2O2 ASR with a gradient of H2O2 concentrations for wt, msn4∆ and skn7∆.

_**Data**_

Species       Strain           Genotype     H2O2         
--------      -------------    ----------   ------------
C. glabrata   yH001, yH002     wildtype    60, 80, 100 mM 
C. glabrata   yH396, yH397     _msn4∆_     20, 40, 60 mM 
C. glabrata   yH422, yH423     _skn7∆_     10, 20, 30, 40 mM 

Five replicates (at three [H2O2]) for wt and _msn4∆_, four replicates (at three [H2O2]) for _skn7∆_.

### Data reformatting
Reformat JY's data into a tidy format. Note that the following data file contains the same information as in the `raw` object imported at the beginning of this rmarkdown.


```{r}
tmp1 <- read_tsv("../input/msn4orskn7_ko_ASR_raw.tsv", col_types = cols(), comment = "#") %>% 
  mutate(Date = gsub("(\\d\\d)(\\d\\d)(\\d\\d)", "\\1/\\2/\\3", Date)) %>% 
  mutate(H2O2 = recode(H2O2, 'Mock' = '0mM')) %>% 
  mutate(H2O2 = gsub("mM", " mM", H2O2)) %>% 
  select(-`MO/MM_percent`, -`PO/PM_percent`)

# data sanity check, quick view
sapply(select(tmp1, Date, Species, Strain, Genotype, Len_1, Len_2, H2O2), unique)

dat.sf <- tmp1 %>% 
  pivot_longer(c(Count_MO_MM, Count_PO_PM), names_to = "Primary", values_to = "Count") %>% 
  mutate(Primary = fct_recode(Primary, "Mock" = "Count_MO_MM", "-Pi" = "Count_PO_PM"),
         scaled = Count * Dilutions * 1e-3) %>% 
  group_by(Date, Strain, Genotype, Primary) %>%
  mutate(scaled_M = scaled[H2O2 == "0 mM"],
         r = scaled / scaled_M) %>%
  ungroup() %>% 
  select(Date, Species, Strain, Genotype, Primary, H2O2, Dilutions, 
         Count, scaled, scaled_M, r) %>% 
  arrange(Date, Genotype, Primary)
```
Save the data for future references
```{r}
write_tsv(dat.sf, file = "../input/20230913-sf-data-JY.tsv")
```

### Plot (all data)
filter out skn7Δ group in 40mM treatment, which is too high according to the spotting assay result. 

```{r}
dat.sfp <- dat.sf %>% 
  filter(H2O2 != "0 mM", !(H2O2 == "40 mM" & Genotype == "skn7Δ")) %>% 
  mutate(
    Genotype = factor(Genotype, levels = c("wt", "msn4Δ", "skn7Δ")),
    group = paste(Genotype, gsub(" ", "", H2O2), sep = "_"),
    Secondary = factor(group, 
                       levels = c("wt_100mM", "wt_80mM", "wt_60mM", "msn4Δ_60mM", "msn4Δ_40mM", "msn4Δ_20mM",
                                   "skn7Δ_30mM", "skn7Δ_20mM", "skn7Δ_10mM"),
                       labels = c("High", "Medium", "low", "High", "Medium", "low", "Medium", "low", "bottom"))
  )
with(dat.sfp, table(Date, group))
```

```{r}
dat.sfp %>%
  ggplot(aes(x = H2O2, y = r)) + #p.survival[-3] +
  geom_point(aes(shape = Primary), stroke = 1, size = 2, 
             position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.9)) +
  stat_summary(aes(group = Primary), position = position_dodge(0.9),
               fun = mean, fun.max = mean, fun.min = mean,
               geom = "crossbar", color = "red", width = 0.5) +
  facet_wrap(~ Secondary + Genotype, nrow = 1, scales = "free_x") +
  scale_shape_manual(values = c("Mock" = 1, "-Pi" = 16)) +
  scale_y_continuous(labels = scales::percent) +
  xlab(bquote(H[2]*O[2]~(mM))) + ylab("% survival") +
  theme_bw(base_size = 14, base_line_size = 1) +
  panel_border(color = "black", size = 1) +
  theme(strip.text = element_text(size = rel(1), face = 3),
        strip.background = element_blank())
```

### Basal survival rates

Statistical test for differences in basal survival between genotypes. I first applied the Kruskal-Wallis test, which is an extension of the Mann-Whitney's U test for multiple groups and is equivalent to ANOVA but applied on ranks. I used this test to determine if there is evidence for differences in the basal survival rates among the four groups (2 genotypes x 2 [H2O2]).

```{r}
tmp <- dat.sfp %>% filter(Primary == "Mock")
tmp %>% group_by(group) %>% 
  summarize(
    test = Hmisc::smean.cl.normal(r) %>% t() %>% as_tibble()
  ) %>% 
  unnest(test) %>% 
  mutate(across(where(is.double), round, digits = 3))
kruskal.test(r ~ group, data = tmp)
```
> There are differences among the four groups. Inspecting the statistical summary, we can see that the high and low concentrations used for each genotype seem to match each other well.
> Therefore, we will next use A Wilcoxon rank-sum test to compare the two genotypes at either the high or the low dose

Statistical test for differences in basal survival within genotypes at 80 (wt) / 40 (msn4∆) / 30 mM (skn7∆).
```{r}
tmp1 <- dat.sfp %>% 
  filter(group %in% c("wt_80mM", "msn4Δ_40mM"), Primary == "Mock")
wilcox.test(r ~ group, data = tmp1, paired = FALSE)

tmp1 <- dat.sfp %>% 
  filter(group %in% c("wt_80mM", "skn7Δ_30mM"), Primary == "Mock")
wilcox.test(r ~ group, data = tmp1, paired = FALSE)

tmp1 <- dat.sfp %>% 
  filter(group %in% c("msn4Δ_40mM", "skn7Δ_30mM"), Primary == "Mock")
wilcox.test(r ~ group, data = tmp1, paired = FALSE)
```
> We can conclude that there is no significant difference in survival when comparing 80 mM vs 40 mM vs 30mM for wt and msn4∆ and skn7∆.

### 80 vs 40 vs 30 mM
#### Plot
```{r}
dat.sfp %>% filter(Secondary == "Medium") %>% 
  mutate(Genotype = fct_recode(Genotype, `WT` = "wt")) %>% 
  ggplot(aes(x = Primary, y = r)) + 
  #geom_line(aes(group = paste0(Date, Strain, Secondary)),
  #          linetype = 2, linewidth = 0.1) + 
  geom_point(aes(shape = Primary), stroke = 1, size = 2,
             position = position_jitter(0.05)) +
  scale_shape_manual(values = c(1, 16), guide = "none") +
  p.survival[-1] +
  facet_wrap(~Genotype + H2O2, nrow = 1) +
  panel_border(color = "black", size = 1.5) +
  theme(axis.line = element_blank(),
        axis.title.x = element_blank(), 
        strip.background = element_blank(),
        legend.position = "top",
        legend.justification = "center",
        legend.text = element_text(size = rel(0.9)),
        legend.title = element_text(size = rel(0.9)))
#ggsave("../output/20230522-rim15-vs-wt-ASR.png", width = 3.6, height = 3.4)
```
#### Statistical test
1. Determine if the primary stress enhanced the survival in _each_ genotype (Wilcoxon signed-rank test)

_**ASR effect size between wt and msn4∆**_
```{r}
tmp <- dat.sfp %>% 
  filter(Secondary == "Medium", !(Strain %in% c("yH422", "yH423"))) %>% 
  pivot_wider(id_cols = c(Date, Strain, Genotype), names_from = Primary, values_from = r) %>% 
  mutate(ASR = `-Pi`/Mock) %>% 
  arrange(Genotype)

tmp %>% group_by(Genotype) %>% 
  summarize(ASR_score = paste(round(as.numeric(ASR),1), collapse = ", "), 
            mean = mean(ASR), sd = sd(ASR))

t.test(as.numeric(ASR) ~ Genotype, paired = FALSE, data = tmp)
wilcox.test(as.numeric(ASR) ~ Genotype, paired = FALSE, data = tmp)
```
> No significant difference detected in the ASR-score between _msn4∆_ and wild type is significant at a 0.05 level (P = 0.53).


_**ASR effect size between wt and skn7∆**_
```{r}
tmp <- dat.sfp %>% 
  filter(Secondary == "Medium", !(Strain %in% c("yH396", "yH397"))) %>% 
  pivot_wider(id_cols = c(Date, Strain, Genotype), names_from = Primary, values_from = r) %>% 
  mutate(ASR = `-Pi`/Mock) %>% 
  arrange(Genotype)

tmp %>% group_by(Genotype) %>% 
  summarize(ASR_score = paste(round(as.numeric(ASR),1), collapse = ", "), 
            mean = mean(ASR), sd = sd(ASR))

t.test(as.numeric(ASR) ~ Genotype, paired = FALSE, data = tmp)
wilcox.test(as.numeric(ASR) ~ Genotype, paired = FALSE, data = tmp)
```
> No significant difference detected in the ASR-score between _skn7∆_ and wild type is significant at a 0.05 level (P = 0.17).








### Old Code Chunk


```{r}
asr_raw <- read.csv("../input/MOPO_Sc_Cg_ASR_scatter_plot_d122422.csv")

# scatterplot of MO/MM vs PO/PM
#png("../output/MO_vs_PO.png", width = 4, height = 4, units = 'in',res = 300) # png for slides
asr_raw  %>% subset(., MO.MM != 1 & PO.PM != 1) %>%
  ggplot(aes(x = MO.MM, y = PO.PM)) + 
  geom_point(aes(color = Strain),size = 2) + 
  geom_smooth(aes(color = Strain)) +  # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_x_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  # change the scale of x and y axis
  coord_fixed(ratio = 1, xlim = c(0,0.4), ylim = c(0,0.4), expand = TRUE, clip = "on") +
  geom_abline(slope = 1, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("PO/PM") +
  xlab("MO/MM") +
  ggtitle("PO/PM vs MO/MM")


# scatterplot of MO/MM vs PO/PM
asr_raw  %>% subset(., MO.MM != 1 & PO.PM != 1) %>%
  ggplot(aes(x = MO.MM, y = PO.PM)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #geom_smooth(aes(color = Strain)) +  # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_x_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  # change the scale of x and y axis
  coord_fixed(ratio = 1, xlim = c(0,0.4), ylim = c(0,0.4), expand = TRUE, clip = "on") +
  geom_abline(slope = 1, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("PO/PM") +
  xlab("MO/MM") +
  ggtitle("PO/PM vs MO/MM")
#  ggsave("../output/MO_vs_PO_HT.png", width = 4, height = 4, units = 'in',dpi = 300)
  
# scatterplot of MO/MM vs PO/PM
asr_raw  %>% subset(., Strain != "Kl" & MO.MM != 1 & PO.PM != 1) %>%
  ggplot(aes(x = MO.MM, y = PO.PM)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #geom_smooth(aes(color = Strain)) +  # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_x_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  # change the scale of x and y axis
  coord_fixed(ratio = 1, xlim = c(0,0.4), ylim = c(0,0.4), expand = TRUE, clip = "on") +
  geom_abline(slope = 1, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("PO/PM") +
  xlab("MO/MM") +
  ggtitle("PO/PM vs MO/MM")
#  ggsave("../output/MO_vs_PO_ScCg_HT_d122422.png", width = 4, height = 4, units = 'in',dpi = 300)
  
# scatterplot of MO/MM vs PO/PM
asr_raw  %>% subset(., Strain != "Kl" & Experimenter == "JL" & MO.MM != 1 & PO.PM != 1) %>%
  ggplot(aes(x = MO.MM, y = PO.PM)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #geom_smooth(aes(color = Strain)) +  # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_x_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  # change the scale of x and y axis
  coord_fixed(ratio = 1, xlim = c(0,0.4), ylim = c(0,0.4), expand = TRUE, clip = "on") +
  geom_abline(slope = 1, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("PO/PM") +
  xlab("MO/MM") +
  ggtitle("PO/PM vs MO/MM")
#  ggsave("../output/d122422_MO_vs_PO.png", width = 4, height = 4, units = 'in',dpi = 300)

```

```{r}
# MO/MM vs ASR Score
asr_raw %>% mutate(.,ASR_Score = PO.PM/MO.MM, .keep = "all") %>% subset(., Strain != "Kl" & MO.MM != 1 & ASR_Score != 1) %>%
  ggplot(aes(x = MO.MM, y = ASR_Score)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_x_continuous(labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  # change the scale of x and y axis
  coord_fixed(ratio = 0.01, xlim = c(0,0.4), ylim = c(0,40), expand = TRUE, clip = "on") +
  geom_abline(slope = 0, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("ASR_Score") +
  xlab("MO/MM") +
  ggtitle("ASR effect vs MO/MM")
#  ggsave("../output/ASR_vs_MOvsMM.png", width = 4, height = 4, units = 'in',dpi = 300)

# MO/MM vs ASR Score with loess fit
asr_raw %>% mutate(.,ASR_Score = PO.PM/MO.MM, .keep = "all") %>% subset(.,Strain != "Kl") %>% subset(., MO.MM != 1 & ASR_Score != 1) %>%
  ggplot(aes(x = MO.MM, y = ASR_Score)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #geom_smooth(aes(color = Strain),span = 1) + # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(trans = "log10")+
  scale_x_reverse(lim = c(0.3,0), labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  #scale_x_reverse() +
  # change the scale of x and y axis
  #coord_fixed(ratio = 0.01, xlim = c(0,0.4), ylim = c(0,40), expand = TRUE, clip = "on") +
  coord_fixed(ratio = 0.1,xlim = c(0.3,0),ylim = c(0.3,30)) +
  geom_abline(slope = 0, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("ASR_Score") +
  xlab("MO/MM") +
  ggtitle("ASR effect vs MO/MM")
#  ggsave("../output/ASR_vs_MOvsMM_rev_x_logy.png", width = 4, height = 4, units = 'in',dpi = 300)
  
# MO/MM vs ASR Score with loess fit
asr_raw %>% mutate(.,ASR_Score = PO.PM/MO.MM, .keep = "all") %>% subset(., MO.MM != 1 & ASR_Score != 1) %>%
  ggplot(aes(x = MO.MM, y = ASR_Score)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #geom_smooth(aes(color = Strain),span = 1) + # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(trans = "log10")+
  scale_x_reverse(lim = c(0.3,0), labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  #scale_x_reverse() +
  # change the scale of x and y axis
  #coord_fixed(ratio = 0.01, xlim = c(0,0.4), ylim = c(0,40), expand = TRUE, clip = "on") +
  coord_fixed(ratio = 0.1,xlim = c(0.3,0),ylim = c(0.3,30)) +
  geom_abline(slope = 0, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("ASR_Score") +
  xlab("MO/MM") +
  ggtitle("ASR effect vs MO/MM")
#  ggsave("../output/ASR_vs_MOvsMM_rev_x_logy_KL.png", width = 4, height = 4, units = 'in',dpi = 300)

# MO/MM vs ASR Score with loess fit
asr_raw %>% mutate(.,ASR_Score = PO.PM/MO.MM, .keep = "all") %>% subset(., Strain != "Kl" & MO.MM != 1 & ASR_Score != 1) %>%
  ggplot(aes(x = MO.MM, y = ASR_Score)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #geom_smooth(aes(color = Strain),span = 1) + # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(trans = "log10")+
  scale_x_reverse(lim = c(0.3,0), labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  #scale_x_reverse() +
  # change the scale of x and y axis
  #coord_fixed(ratio = 0.01, xlim = c(0,0.4), ylim = c(0,40), expand = TRUE, clip = "on") +
  coord_fixed(ratio = 0.1,xlim = c(0.3,0),ylim = c(0.3,30)) +
  geom_abline(slope = 0, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("ASR_Score") +
  xlab("MO/MM") +
  ggtitle("ASR effect vs MO/MM")
#  ggsave("../output/ASR_vs_MOvsMM_rev_x_logy.png", width = 4, height = 4, units = 'in',dpi = 300)

# MO/MM vs ASR Score with loess fit
asr_raw %>% mutate(.,ASR_Score = PO.PM/MO.MM, .keep = "all") %>% subset(., Experimenter == "JL", Strain != "Kl" & MO.MM != 1 & ASR_Score != 1) %>%
  ggplot(aes(x = MO.MM, y = ASR_Score)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #geom_smooth(aes(color = Strain),span = 1) + # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(trans = "log10")+
  scale_x_reverse(lim = c(0.3,0), labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  #scale_x_reverse() +
  # change the scale of x and y axis
  #coord_fixed(ratio = 0.01, xlim = c(0,0.4), ylim = c(0,40), expand = TRUE, clip = "on") +
  coord_fixed(ratio = 0.1,xlim = c(0.3,0),ylim = c(0.3,30)) +
  geom_abline(slope = 0, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("ASR_Score") +
  xlab("MO/MM") +
  ggtitle("ASR effect vs MO/MM")
#  ggsave("../output/d122422_ASR_vs_MOvsMM_rev_x_logy_JL.png", width = 4, height = 4, units = 'in',dpi = 300)


# MO/MM vs ASR Score for Cg
asr_raw %>% subset(.,Strain == "Cg") %>% mutate(.,ASR_Score = PO.PM/MO.MM, .keep = "all") %>% subset(., MO.MM != 1 & ASR_Score != 1) %>%
  ggplot(aes(x = MO.MM, y = ASR_Score)) + 
  geom_point(aes(color = Strain),size = 2) + 
  #geom_smooth(aes(color = Strain),span = 1) + # fit to a loess model with 95% CI 
  #geom_smooth(aes(color = Strain, fill = Strain), method = "lm") +
  #scale_color_viridis(discrete = TRUE, option = "D") +
  #scale_fill_viridis(discrete = TRUE) + 

  # change decimal scale to pct
  scale_y_continuous(trans = "log10")+
  scale_x_reverse(lim = c(0.3,0), labels = scales::percent, breaks = seq(0, 0.4, 0.1)) +
  scale_color_manual(values = c("grey", "deeppink3", "hotpink", "darkorchid2", "deepskyblue", "dodgerblue3", "grey20", "seagreen1"))  +
  #scale_x_reverse() +
  # change the scale of x and y axis
  #coord_fixed(ratio = 0.01, xlim = c(0,0.4), ylim = c(0,40), expand = TRUE, clip = "on") +
  coord_fixed(ratio = 0.1,xlim = c(0.3,0),ylim = c(0.3,30)) +
  geom_abline(slope = 0, intercept = 0, color="wheat2", linetype="dashed", size=1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("ASR_Score") +
  xlab("Severity of the secondary stress") +
  ggtitle("ASR effect vs MO/MM")
 # ggsave("../output/Cg_ASR_vs_MOvsMM_rev_x_logy.png", width = 4, height = 4, units = 'in',dpi = 300)  
  
  
```

# plot the comparable MO data at 100mM and 10mM for Cg and Sc
# do the statistical test
```{r}
mo_comp <- asr_raw %>% filter(H2O2 %in% c("100mM","10mM"))
mo_comp %>% ggplot(aes(x = Strain, y = MO.MM_percent)) + geom_point(size =5, alpha = 0.5) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("Survival Rate % (MO/MM)") +
  xlab("Strain")
# ggsave("../output/Cg100_Sc10.png", width = 4, height = 4, units = 'in',dpi = 300) 


# test for homogeneity in variance
res.ftest <- var.test(MO.MM_percent ~ Strain, data = mo_comp)
res.ftest
## There is no significant difference between the variances of the two sets of data. Therefore, we can use the classic t-test witch assume equality of the two variances.

# paired t-test
res <- t.test(MO.MM_percent ~ Strain, data = mo_comp, paired = TRUE, var.equal = TRUE)
res
## no significant difference in the survival rate of 100mM treated Cg and 10mM treated Sc
```

# Fold increase in survival relative to the primary stress (CFU data for Cg and Sc)
```{r}
asr_dot <- asr_raw %>% filter(H2O2 %in% c("20mM","40mM","60mM","80mM","100mM","2mM","4mM","6mM","8mM","10mM") & !(Strain %in% c("Kl"))) %>% mutate(H2O2_con = as.numeric(gsub("mM","",H2O2)))

# Cg CFU
asr_dot %>% filter(Strain == "Cg") %>% 
  ggplot(aes(x = H2O2_con, y = ASR_score)) + geom_point(dotsize = 5, alpha = 0.5) + 
  scale_y_continuous(limits = c(0, 33)) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("ASR Score (r'/r)") +
  xlab("H2O2 conc. (mM)")
 ggsave("../output/Cg_ASR_grad_CFU.png", width = 4, height = 4, units = 'in',dpi = 300) 

# Sc CFU
asr_dot %>% filter(Strain == "Sc") %>% 
  ggplot(aes(x = H2O2_con, y = ASR_score)) + geom_point(dotsize = 5, alpha = 0.5) + 
  scale_y_continuous(limits = c(0, 33)) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("ASR Score (r'/r)") +
  xlab("H2O2 conc. (mM)")
 ggsave("../output/Sc_ASR_grad_CFU.png", width = 4, height = 4, units = 'in',dpi = 300) 

# combine the Cg and Sc onto the same plot
asr_dot %>%
  ggplot(aes(x = H2O2_con, y = ASR_score)) + geom_point(aes(color = Strain),dotsize = 5, alpha = 0.5) + 
  scale_y_continuous(limits = c(0, 33)) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("ASR Score (r'/r)") +
  xlab("H2O2 conc. (mM)")

```

# Fold increase in survival relative to the primary stress (spotting assay density data for Cg and Sc)
```{r}
spot_asr <-  read.csv("../input/spotting_assay_density.csv")

# Cg spotting assay
spot_asr %>% filter(Strain == "Cg") %>% 
  ggplot(aes(x = H2O2_mM, y = ASR_Score)) + geom_point(dotsize = 3, alpha = 0.5) +  
  scale_y_continuous(limits = c(0, 60)) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("ASR Score (r'/r)") +
  xlab("H2O2 conc. (mM)")
 ggsave("../output/Cg_ASR_grad_spot.png", width = 4, height = 4, units = 'in',dpi = 300) 

# Sc spotting
spot_asr %>% filter(Strain == "Sc") %>% 
  ggplot(aes(x = H2O2_mM, y = ASR_Score)) + geom_point(dotsize = 3, alpha = 0.5) +  
  scale_y_continuous(limits = c(0, 10)) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("ASR Score (r'/r)") +
  xlab("H2O2 conc. (mM)")
 ggsave("../output/Sc_ASR_grad_spot.png", width = 4, height = 4, units = 'in',dpi = 300) 


# combine Cg and Sc
spot_asr %>% ggplot(aes(x = H2O2_mM, y = ASR_Score)) + geom_point(aes(color = Strain), dotsize = 3, alpha = 0.5) + 
  scale_y_continuous(limits = c(0, 60)) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("ASR Score (r'/r)") +
  xlab("H2O2 conc. (mM)")

  

```

# ASR Score vs different length of the primary stress treatment (CFU data)
```{r}
time_asr <- read.csv("../input/Time_course_ASR.csv")

time_asr %>%  ggplot(aes(x = Time_min, y = ASR_Score)) + geom_point(aes(color = Strain), dotsize = 5, alpha = 0.5) +
    scale_x_continuous(minor_breaks = seq(0,135,by=45),
                     breaks = seq(0,135,by=45),
                     limits = c(40,140),
                     expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 250)) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("ASR Score (r'/r)") +
  xlab("Time (min)")
 ggsave("../output/timecourse_asr.png", width = 4, height = 4, units = 'in',dpi = 300) 

# transformed the y axis
time_asr %>%  ggplot(aes(x = Time_min, y = ASR_Score)) + geom_point(aes(color = Strain), dotsize = 5, alpha = 0.5) +
    scale_x_continuous(minor_breaks = seq(0,135,by=45),
                     breaks = seq(0,135,by=45),
                     limits = c(40,140),
                     expand = c(0,0)) +
  scale_y_continuous(trans = 'log10',limits = c(0.1, 250)) +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1)) +
  ylab("ASR Score (r'/r)") +
  xlab("Time (min)")

```



